---
output: github_document
---

# Rcpp + OpenMP

This repository shows how much speed gains can be obtained from using
[OpenMP](https://openmp.org), and in particular, the `omp simd` and `parallel for`
instructions.

The test consists on computing the pair-wise distances between rows in a matrix
of size `N` by `M`. The equivalent function in R is `dist()`, and here we 
redefined it using [`Rcpp`](https://cran.r-project.org/package=Rcpp). The program is

The file [norm.cpp](norm.cpp) contains the C++ source code for the dist functions.

## Speed benchmark

Whe using multicore, I'm only using 2 cores

```{r execution, cache=TRUE}
Rcpp::sourceCpp("norm.cpp")

library(microbenchmark)
set.seed(718243)
N <- 200
M <- 1000
x <- matrix(runif(N * M), nrow = N)

(ans_bm <- microbenchmark(
  `SIMD + parfor` = dist_omp_simd(x, N, M, 2),
  `parfor`        = dist_omp(x, N, M, 2),
  `SIMD`          = dist_simd(x, N, M),
  `serial`        = dist_for(x, N, M),
  R               = dist(x),
  times           = 200,
  unit            = "relative"
))
```

Overall, in my machine, the SIMD+parfor combo outperforms all the others. Let's
see if the results are equivalent. At the very least, we should only observe
small differences (if any) b/c of precision:

```{r Comparing-results, cache=TRUE}
Rcpp::sourceCpp("norm.cpp")
ans0 <- as.matrix(dist(x))
ans1a <- dist_omp_simd(x, N, M)
ans1b <- dist_omp(x, N, M)
ans1c <- dist_simd(x, N, M)
ans1d <- dist_for(x, N, M)
range(ans0 - ans1b)
range(ans1a - ans1b)
range(ans1b - ans1c)
range(ans1c - ans1d)
```


Finally, the [`microbenchmark`](https://cran.r-project.org/package=microbenchmark)
R package offers a nice viz with boxplot comparing all the methods:

```{r viz, dependson='execution'}
op <- par(mai = par("mai") * c(2,1,1,1))
boxplot(ans_bm, las = 2, xlab = "")
par(op)
```

## Session info

```{r}
sessionInfo()
```

